\documentclass[main.tex]{subfiles}
\begin{document}

\chapter{The Beast Awakens...}

\par \nar \rmelectra~ continued to wander closer to the very center of the Milky Way's nuclear star cluster.  The stars in her immediate vicinity seemed to fly by faster and faster as she drew ever closer to her destination.  Such high relative velocities could \textit{only} be imparted via the gravitational acceleration supplied by some very nearby compact massive object, a description that fits \rmchiron~ to a ``T''.  

\par \nar Random perturbations from passing stars continuously deflected her orbit, causing her to wander throughout the nuclear star cluster somewhat aimlessly for what seemed like a very longtime.  All the while, she searched for some sign of \rmchiron.  Until one day, she saw the tell-tale signs of a thin, fleeting accretion disk.  A soft halo of light surrounded a black sphere.  \rmenrico~ had proved right:  the super-massive black hole was indeed uniquely identified by the thin accretion disk \rmenrico~ had described to her, formed of the coalesced winds of nearby massive stars.\footnote{When matters falls deep into the local gravitational potential of a massive compact object, it often gathers into an "accretion disk".  This is because the infalling matter carries finite angular momentum, which must be conserved.  Hence, close to the black hole, the gas particles will sort themselves into an approximately planar configuration with the orbital plane being aligned with the total angular momentum vector.  This planar configuration of gas particles is called an "accretion disk", since viscosity within the disk can cause angular momentum to be redistributed and dissipated.  Hence, the inner material of the disk can lose enough angular momentum this way to eventually be accreted directly onto the compact object.  The gas particles in the disk collide with each other, emitting light in the process and making accretion disks visible to observers on Earth.}

\section{The Behemoth of Black Holes}

\par \Electra  Hello there!  If I squint, I see what looks like an accretion disk hiding over there.  Is that you \rmchiron?

par \nar A long silence followed \rmelectra~'s seemingly innocent question.  Eventually, the silence was broken. A deep booming voice responded, shaking \rmelectra~ to her core.

\par \Chiron Yes.  It is I, \rmchiron.  I am here, and you are there.  But who is it that I am speaking to?

\par \Electra I am \rmelectra~.  A wandering star in search of my lost sisters.

\par \Chiron I see.  You have spotted what remains of the corpse of my last meal.  The scene before you might look a little gruesome but, I assure you, it is but one aspect of all that must be.  I do my part, here in this most extreme of environments in our Galaxy.  The stellar density can only grow so high, after all.  Otherwise, we'd all be banging into each other on a nearly daily basis.\footnote{Recall that the stellar densities can reach up $> 10^7$ M$_{\odot}$ pc$^{-3}$ in the Milky Way's central nuclear cluster.  At such high densities, the rate of direct collisions between single and binary stars, and even between two single stars, can be very high.  There is also the issue of stars diffusing into the "loss cone" of the central SMBH, which refers to the family of orbits that drift sufficiently close to the SMBH that the orbiting star will be eaten by the SMBH, usually in the form of a tidal disruption event (see below).  This also serves to deplete stars from the local stellar population, reducing the stellar density.  The diffusion rate of stars into the SMBH loss cone be very high when the density is high, but drops off significantly at low densities.} 

\par \Electra Of course you do!  I did not mean to accuse you of anything.  I cannot thank you enough for all that you do.  I am here in the hopes of meeting you, to draw upon your incredible wisdom and learn about my place in the grand scheme of things.  I was born of a particularly compelling Giant Molecular Cloud, and have been wandering my way through the Galaxy ever since, independent of my siblings.  As did I, they ventured off in random directions when my Mother dispersed.

\par \Chiron I see.  A quest!  But with what purpose?

\par \Electra That is the question!  I suppose I am here to understand.  To fully digest this Universe we call home, while avoiding falling victim to it.  Experience its many wonders and breath-taking panaromas!  Eventually, I hope to reunite with my sisters, and hear their lovely voices one more time.

\par \Chiron.  A good answer, young one.  You have come to the right place, if gaining perspective is your goal.  This part of the Galaxy is unlike any other.  Extreme stellar densities, recent and even ongoing star formation, collisions and high-velocity impacts mediated by a super-massive black hole (i.e., me), etc.\footnote{If a star travels very close to an SMBH, a significant gravitational acceleration can be imparted.  in fact, if another star or BH is involved (required for conservation of energy and angular momentum), then a "hypervelocity star" (HVS) can be produced, escaping with a final velocity up to many thousands of kilometers per second (provided it avoids a tidal disruption event; see above).  Such HVSs have actually been observed in our Galaxy, with a few candidates seeming to have originated from the Galactic Center, as expected if the HVSs formed from an interaction with the central SMBH.}  Being here and witnessing these types of events will surely help you to better understand them, gaining further insight to the inner machinery of the Universe.

\par \nar Wide-eyed and eager for answers, \rmelectra~ was all in.
%NL: Does the above make sense to a common readier.  It is sort of jargon...

\par \Electra I am intrigued!  How \textit{do} they work?

\par \Chiron Come a little closer, child.  I can barely hear you.

\par \Electra Oh sure, of course.  That makes sense.

\par \nar \rmelectra~ deflected her orbit to bring her a little closer to \rmchiron.

\par \Chiron The first thing I shall teach you is that, the closer you are to a massive object and, more generally, the more mass you have very close to you, the acceleration imparted to you per unit time will be greater.  In other words, you will notice that your velocity relative to your accelerator and even other nearby stars, increases.

\par \Electra Yes!  I have noticed that!  Just now, I came closer to you, and I am now orbiting you at higher velocity.  Is that the effect you speak of?

\par \Chiron Yes, that is exactly it.  Very good.  But I can still hardly hear you.  Come a little closer still; be kind to these old ears.

\par \nar \rmelectra~ drew ever closer to \rmchiron, her typical orbital velocity now several hundreds of kilometers per second, and her mean orbital distance no more than a fraction of a parsec.  Her orbit about \rmchiron~ had grown increasingly eccentric and, at pericenter, she was about a thousand times her own radius from \rmchiron's center of mass.\footnote{Note that this lies well outside the tidal disruption radius.  Another order-of-magnitude reduction in the distance at closest approach, however, and TDEs should become increasingly common.}  

\par \Chiron That is good.  I can now resolve your face with these old eyes, at least.  Wonderful to meet you, my child.

\par \Electra  And you, \rmchiron!  What else can you tell me about the Universe?

\par \Chiron Well, you have probably noticed that more massive objects exert a greater gravitational acceleration, at a given distance from their center of mass.  

\par \Electra  Oh yes.  That was one of the first things that I noticed.  When my Mother dispersed, and her considerable mass along with it, my siblings and I became gravitationally unbound.  I have not seen them since, but the memory is still with me as if it only just happened yesterday.  I do miss them.

\par \Chiron I can only imagine, child.  Perhaps I can help you to find them.

\par \Electra That would be wonderful!  Do you know where they are?

\par \Chiron I think that I might.  Come a little closer still, child, this is a secret best kept between friends.  We will need to whisper in each other's ears in such a crowded environment.

\par \nar \rmelectra~ did as requested, bringing her pericenter distance even closer to \rmchiron.  But she could now feel the consequences of her decision; at apocenter, her velocity was very low relative to other stars in her vicinity.  This meant that she could feel the individual gravitational perturbations from passing stars; each one seemed to knock her on to an even more eccentric orbit relative to \rmchiron.  If this continued, she would only drift closer to him.  It seemed there was now no going back.\footnote{DESCRIBE HOW SUCH PERTURBATIONS INCREASE ECCENTRICITY, USING CONSERVATION OF ENERGY AND ANGULAR MOMENTUM.}

\par \nar \rmelectra~ realized she had mistaken a gaping maw for a warm smile.  Now that she could actually see \rmchiron's face, she recognized the hunger.  And she suddenly realized she had been selected as the next meal.  But her fate was now sealed.  Without outside help, there was nothing she could do to avoid ever higher eccentricities, and hence ever smaller distances of closest approach.  

\section{A Change of Heart in the Heart of the Milky Way}

\par \nar And, sure enough, a few orbits later, \rmelectra~ came within about a 100 times her radius from \rmchiron's center of mass.  A differential force like she had never felt before was exerted upon her.  She was stretched (along the direction of her orbital motion) and squeezed (perpendicular to her orbital motion) as she completed each orbit, and it was only getting worse as time passed.  In spite of this discomfort, \rmelectra~ remained blissfully unaware of \rmchiron's true intentions.

\par \Electra I do like you, \rmchiron.  You are being so kind to me.  You offer answers, and yet I have nothing to offer you in return.  Is there anything I can do?  A hug, at the very least, perhaps?

\par \nar \rmelectra~ tried to extend herself in \rmchiron's direction, hoping to bestow upon him a firm embrace like he had never known before.  \rmchiron~ could see the sincerity etched across \rmelectra's face, scrunched up in concentration as she tried to deliver the hug she felt strongly \rmchiron~ surely must need.  Touched by this, \rmchiron~ suddenly began to cry.  Deep, woeful sobs racked his torso.

\par \Electra Oh my goodness!  You are in pain!  What is wrong?  Have I done something wrong?

\par \Chiron No, child.  It is quite the opposite.  You have reminded me of the purity and sincerity you young, hopeful stars often possess.  Ah, how I have missed that.  It has been a long time, you must understand.  

\par \Electra Oh, I think I can understand.  I mean....I'm happy to try, at least.

\par \nar sadness appeared suddenly on \rmchiron's face.

\par \Electra What is wrong, \rmchiron?  Please allow me to help.

\par \Chiron I fear there is something about me you \textit{cannot} understand.

\par \Electra Well, try me!  You'll have to give it a go to find out for sure.

\par \nar \rmchiron~ sighed.

\par \Chiron Very well.  I was once like you.  A young star, wandering the Universe.  Completely free and excited for the adventures I was sure would befall me.  And, indeed, I saw many an adventure.  But, like all things, I eventually reached the end of my long life.  I was a massive star, once upon a time, and when my nuclear fuel ran out I became a black hole.  Back then, I wasn't much more massive than you are now, even as a black hole.  But, because of our considerable mass, we black holes are cursed to encounter other objects on much shorter timescales than other stars, especially in dense environments.\footnote{Recall that the rate of direct collisions is proportional to the collisional cross-section, which defines an area that, if both stars pass through it at the same time, their radii would overlap directly and a collision would occur.  Provided objects have sufficiently slow relative velocities, "gravitational focusing" (i.e., gravity curves the relative trajectories of the passing objects and focuses them more toward each other) can significantly increase the collisional cross-section and enhance the collision rate.}  And, when we do, there is a good chance we will consume them, devouring them whole.  Each time, we become a little more massive, slowly growing.  

\par \Chiron In my case, I found myself in a very dense environment when I first became a black hole.  Stars whipped by me all the time.  Before I knew it, they were drifting too close to me, and my gravitational pull ripped them from this world.  They call it a \textit{tidal disruption event}.\footnote{Recall that this is when a star passes too close to an SMBH, and is torn apart by the induced tidal forces, often triggering a very bright flare.}  At closest approach to the black hole, the culprit star is stretched along the direction of its orbit, and squeezed in the direction perpendicular to it.  The star's brightens grows immensely due to this compression, at least briefly, but most of its mass falls back and is ultimately eaten by the black hole.  

\par \Electra So you ate a lot of stars?

\par \Chiron Yes.  It was accidental at first, but eventually I became fixated on the idea of becoming even more massive.  I became mad, and left my \textit{stellarity}\footnote{Stellarity is to stars what humanity is to humans.} behind.  I do not want to be this devouring black hole anymore.  I wish to help and connect with my fellow stars and remnants.  

\par \Electra Well, I forgive you, \rmchiron.  And I appreciate your honesty.

\par \nar Once again, \rmelectra~ struggled, unsuccessfully, to hug to \rmchiron.

\par \Chiron Thank you, child.  I am truly glad to have met one such as you, a kind soul out here in the harshest of environments.

\par \Electra Aw, shucks.  Happy to help.  And I'm glad to have met you too!

\par \Chiron To be honest child, I was begging you ever closer earlier in an attempt to consume you.  But I no longer have the stomach for it.  I am far too fond of you now.

\par \Electra And I you, \rmchiron.  I was wondering why I kind of feel like the progenitor of one of those tidal disruption events you mentioned every time I reach pericenter...

\par \Chiron It is still good news!  From your current location, I can impart a significant acceleration to you.  This would easily give a sufficiently high velocity for you to escape from here, and hopefully find your lost siblings.

\par \Electra  That would be great!  What do we need to do?

\par \Chiron Well, we will need another moderately massive star or remnant.  More massive than you, certainly.  This will be needed to maximize the probability that it is you who is ejected.\footnote{During a chaotic gravitational interaction involving three or more particles, the least massive particle is almost always the first to be ejected.  The reason is simple:  the imparted force is large and so is the resulting acceleration due, respectively, to the larger masses of the other particles and the lower mass of the ejected particle.  The probability of ejection for the lowest mass particle quickly asymptotes to unity as the mass ratio goes above about 10 or so; for very low-mass particles, they are almost guaranteed to be ejected (if they are not first destroyed during a TDE event or collision).}

\par \Electra Wait, what?  I might \textit{not} be the one ejected?  So this is \textit{not} a guarantee?  Uh.. Okay, let's say I am not ejected.  What do you suppose would happen instead?

\par \Chiron Uh... Well, most likely, I would consume you as a tidal disruption event.

\par \Electra WHAT?!  Are you crazy? 

\par \Chiron Well, you have a very large physical size compared to a black hole.  It \textit{significantly} increases the probability of you colliding with another object directly in \textit{any} scenario.\footnote{Recall that a tidal disruption event (TDE) is what occurs when a star passes too close to a massive black hole.  Close to the BH, gravity is sufficiently strong to deform the star (stretching it along the orbit and squeezing it in the radial direction, perpendicular to the (instantaneous) orbital trajectory).  The star can be squeezed sufficiently to brighten it by many orders of magnitude, often resembling a supernova explosion.  The event is so violent, that the star is usually ripped apart, with some of thee debris remaining unbound and escaping from the immediate vicinity of the BH, and the rest either being accreted directly onto the BH or sorting itself out into an accretion disk around the BH.}  Do not worry though.  It is all about choosing the right third object for the job.  

\par \Electra Okay.  What should the third object be?

\par \Chiron Preferably another massive black hole.  We have small physical sizes, but large masses that can impart significant accelerations at larger distances.

\par \Electra Okay, great.  Let's find another massive black hole!

\section{The Gravitational Slingshot}

\par \Electra Wait.... How exactly \textit{do} we find another black hole?

\par \nar \rmchiron~ whistled suddenly, clearly trying to get someone's attention.

\par \Chiron \rmhippe, my daughter, are you there?

\par \nar Only a moment had passed when the response came, loud and closer than \rmelectra~ had anticipated.

\par \Hippe I am here, father.  What is it that you need?

\par \Chiron You assume I do not seek you simply to talk?  To ask you how you are doing?

\par \Hippe Well?  Is that why you have sought me out?

\par \Chiron Uh... Yes!  How are you, my child!?

\par \Hippe I am well, father.  As you know, this is the one place in the Galaxy that food is plentiful.  

\par \nar \rmchiron glances awkwardly at \rmelectra.  An anxious laugh escapes from \rmchiron's lips.    

\par \Chiron Ha ha, yes, well I have put myself on a very strict diet lately.  I am, after all, the fattest black hole in the Galaxy, with the next fattest black hole lying as far as several hundred megaparsecs away.  No need to exacerbate the problem, your mother used to always say.
%NL: Does all the above make sense?  Feel free to edit, if not.

\par \Hippe Uh... Sure, Dad.  Whatever you say.  So why did you \textit{really} call me?

\par \Chiron To say hello, of course.  But, now that you mention it, I could use a small favor.  

\par \nar \rmhippe~ sighed.

\par \Hippe What is it that you need, father?

\par \Chiron I am in need of a gravity assist for our friend \rmelectra~ here.  She needs to be imparted with a sufficiently high velocity kick that she may venture off throughout the Galaxy in search of her siblings.  

\par \Hippe I see, the ol' gravitational slingshot.  That shouldn't be a problem.  Believe it or not, we black holes are often called upon for a gravity assist.  The curse of being massive!   Well, you can certainly count me in.  It is wonderful to meet you, \rmelectra!

\par \Electra And I you, \rmhippe.  I can't thank you and your father enough for your gracious offer to assist me.  To escape this nuclear star cluster and its deep potential well, I am truly at the mercy of your considerable mass.  I am truly grateful for any acceleration you can bestow upon me. 

\par \Hippe It will be a pleasure, child.  Okay, let's get this show on the road.  Let me just get a running start at you...

\par \Electra Uh... A running start...?

\par \Hippe Yeah, yeah, don't sweat it.  I need to smash in to the both of you at sufficiently high velocity to kick you out and get you where you need to go.

\par \Electra Whoa, whoa, whoa!  Maybe we should talk about...

\par \nar But it was too late.  \rmhippe~ had already gotten her start, and was careening toward \rmelectra's and \rmchiron's mutual center of mass (but skewed toward \rmhippe, the less massive of the two black holes) with startling precision. \rmhippe~ kept her relative velocity low, only a few  kilometers per second relative to \rmchiron~ and \rmelectra's mutual center of mass.  \rmelectra~ screamed in terror as her savior came toward them, the binary's mass pulling \rmhippe~ in and accelerating her toward the pair more noticeably at the last minute.\footnote{Recall that, by Liouville's Theorem, the total volume in phase space for a self-gravitating system of particles is conserved.  In practice, what this means is that interactions with a more negative total energy are more likely to accelerate particles to larger velocities over the course of the interaction.  By keeping her velocity low relative to the binary system center of mass, \rmhippe~ minimized the (positive) kinetic energy she brought in, thereby minimizing the total encounter energy (i.e., by making it more negative, due to the contribution to the total energy from the binary orbital energy).}

\par \nar A rather violent and chaotic dance ensued between \rmelectra~ and the two massive black holes.  When only two bodies are gravitationally bound, Kepler's Laws dictate the time evolution of the system.\footnote{In a nutshell, thanks to Kepler, we can easily calculate everything we would ever want to know for the two-body problem.}  But when a third body is randomly thrown in to the mix, chaos ensues.  All Hell breaks loose.  Each star undergoes a strong gravitational interaction with one or both of the other stars in the system, and recedes on some orbit defined by the exchange of energy and angular momentum during said close interaction.  After completing this temporary orbit, the star returns toward the center of mass of the system.  Upon getting there, each body is perturbed again, and leaves on a new temporary orbit.  This process continues until, eventually, one of the particles is imparted with sufficient kinetic energy that it escapes to spatial infinity.\footnote{This assumes point particles, whereas if finite particle sizes are considered collisions must also be considered}  Due to energy equipartition and conservation of momentum, it is typically the lowest mass star that is ejected, and it usually leaves at the highest possible velocities.  

\par \nar The particular three-body interaction currently under our scrutiny proves to be no exception.  \rmchiron, by far the most massive of the trio by several orders of magnitude, hardly felt the gravitational influence of either \rmelectra~ or \rmhippe.  He moved about little, his velocity relative to the system center of mass never exceeding a few kilometers per second.  His daughter, however, was much less massive than him, and his gravitational influence pushed and pulled her in every which direction.  \rmelectra, by far the least massive of the trio, felt the brunt of it.  She was accelerated more than the other two, orders of magnitude more relative to at least \rmchiron.  Eventually, she escaped at about 1,100 kilometers per second.  But, by some miracle, she was left completely unscathed.%\footnote{REVIEW THE CONCEPTS DRAWN UPON HERE, INCLUDING THE MAXIMUM EJECTION VELOCITY.}

\par \Electra Thank you so much, \rmchiron~ and \rmhippe!  I cannot thank you enough.  I will remember you always, and keep your memory close in my heart.

\par \nar Both \rmchiron~ and \rmhippe~ smiled.

\par \Hippe I see why you liked this one, father.

\par \Chiron She is indeed special.  Innocent and sincere in the most wonderful of ways.

\par \nar \rmhippe~ shouted one last goodbye to her new friend:

\par \Hippe Goodbye and good luck, \rmelectra.  My father was right about you!  You have a warm heart!  I wish you all the luck in the Universe!

\par \nar \rmchiron~ was quick to join suit in extending his best wishes, and goodbyes.

\par \Chiron As do I, child.  Stay true to your heart, and you will be just fine.  May the Cosmos protect you always, and carry you wherever you need to go.

\par \nar Quickly receding in to the distance, \rmelectra~ shouted one final goodbye:

\par \Electra Farewell, to the both of you!  I cannot thank you enough for your help.  I will carry you with me always, and hope you will do the same.  And I will of course tell my siblings all about you, when I find them!

\par \nar \rmelectra~ had now drifted far enough away from the central nuclear star cluster in the Milky Way that she was no longer visible to her new friends; if she could no longer see them, they certainly couldn't see her.  Alone, she was not discouraged.  After all, she was on her way, in search of her lost siblings.  Hopeful that she would soon find them, \rmelectra~ dozed off, falling fast sleep for the next several million years.

%the side of her closest to the black hole was pulled by gravity more than the far side, and she could feel herself being pulled apart.

%SMBH says gluttony robbed him of his hopes and dreams.  He recommends moderation to younger, less massive stellar-mass BHs, so that they may avoid self-consuming themselves before figuring out how to, in a stable way, integrate themselves in to a healthy stellar population.                                    
%Once you feed the beast without limit, the beast will never again know such limits.  Consumption, and more of it.  It must be stopped at all costs and preferably, re-directed to a healthier, more stable fate amidst other stars.  

\end{document}
