\documentclass[main.tex]{subfiles}
\begin{document}

\chapter{Getting to Know your Newest Sibling...}

\par \nar Ever since $\Taygete$ and $\Alcyone$ had merged, life was pleasant for $\Celaeno$.  Peaceful.  $\Celaeno$ and $\Lacedaemon$ had been getting along splendidly.  They became the closest of siblings over the next billion years or so, sharing everything together.  During this time, they had drifted out of the Galactic Plane and were slowly making their way toward the Galactic Halo, now about a kiloparsec above the midplane of the Galactic Disk.\footnote{REFER TO MAP AARON IS GOING TO MAKE!!!} 
%Over the few billion years that followed, $\Celaeno$ and $\Lacedaemon$ continued to drift closer as sisters, while as a pair venturing vertically out of the Galactic Disk.  
Eventually, they found themselves among mostly old stars. What they saw from their new vantage point surprised them; they could now resolve distant stars way out in the Galactic Halo, sufficiently isolated from their peers to escape the bulk of their obscuring light.\footnote{The stellar densities are much higher in the Galactic Disk than in the Galactic Halo.  The more stars that fall along the line of sight, the more light pollution that must be accommodated for (i.e., more of their light must be removed if an observer wishes to see a distant object and measure its brightness reliably).  For example, in city centers, very few, if any, stars tend to be visible in the night sky.  This is largely due to the surrounding light pollution coming from nearby skycrapers, streetlights, car lights, etc., and the additional light makes it difficult to isolate those photons coming from a distant astronomical object of interest, such as a star in a constellation.}  

\par \Celaeno I would recognize our sisters anywhere from all the way out here, given their specific combinations of color and brightness!  Maybe I can spot them! 

\par \nar And, sure enough, $\Celaeno$ saw just such a familiar combination.  Off in the distance, $\Celaeno$ could see a point-source of light that appeared to be getting brighter with each passing year.  For contrast and perspective, a distant cluster within $\Celaeno$'s field of view remained at its initial distant location, keeping the same brightness and angular size on the sky.  Eventually, $\Celaeno$ could resolve the incoming star.  Knowing her sisters' intrinsic brightnesses from having been born right next to them in the same cluster, combined with the observed brightness of the far-off star, $\Celaeno$ was able to calculate a rough estimate as to how far away her distant sister must be, if the distant star is indeed one of her lomng-lost sisters.\footnote{EXPLAIN HOW THIS IS DONE!!!  The intrinsic brightness of a star refers to the star's luminosity as measured at its surface.  Conversely, the apparent brightness of a star depends fundamentally not only on its intrinsic brightness, but also on its distance from the observer. MORE!}

\par \Celaeno  By my calculations, assuming that bright reddish star over there is one of our long lost sisters, it is only a few kiloparsecs away from us.  It is heading right for us.  And fast!  

\par \Lacedaemon I see it!

\par \Celaeno I'd know that combination of color and intrinsic brightness anywhere; unless I miss my guess, that is my sister $\Sterope$!  

\par \nar The millenia passed, and $\Sterope$ drifted ever closer toward $\Celaeno$ and $\Lacedaemon$.  Before she came within ear shot, another sister appeared in the distance.  Apparently coming from the Galactic Center, $\Celaeno$ recognized yet another familiar combination of color and brightness:  $\Electra$.  $\Celaeno$ figured, given a rough guess for her current distance and velocity, $\Electra$ would be upon them within less than a billion years.  This timing worked out nicely with $\Sterope$'s current trajectory.  It seemed a family reunion might truly happen!\footnote{DISCUSS WHY THE PROBABILITY OF THIS IS MINISCULE, AND CALCULATE A ROUGH ESTIMATE FOR THIS PROBABILITY.}

\par \nar Just then, an highly evolved binary star system hosting a millisecond pulsar drifted within sight.  The binary pair were moving very slowly relative to $\Celaeno$ and $\Lacedaemon$; it took another several hundred million years before they came within ear shot.  $\Merope$ and $\Maia$ yelled at their far-off sisters, hoping they would be heard.

\par \Merope Hello there!  I cannot believe what I am seeing!  My siblings!  \textit{Our} siblings!  Look, $\Maia$, our family has returned to us!  

\par \Maia We are beyond ecstatic to see you again in this Universe sisters, let alone the Galaxy of our birth!\footnote{MAYBE DISCUSS THE UNFATHOMABLE TIME IT WOULD TAKE FOR A STAR TO TRAVEL TO ANDROMDEA, OUR CLOSEST NEIGHBOR, AND THE VELOCITY A STAR WOULD NEED TO MAKE IT WITHIN A HUBBLE TIME.}

\par \nar $\Celaeno$ looked intently at $\Merope$ and $\Maia$, not recognizing them at first.  She stared, squinted, focused and concentrated.  Nothing.  A few thousand more years passed; $\Merope$ and $\Maia$ drifted ever closer.  Eventually, recognition spread across $\Celaeno$'s visage.

\par \Celaeno My sister, $\Merope$, is that you?  My old eyes can hardly resolve you.  You are so svelte now!  You must have shed over ten times your original weight when last you left us.  And, oh my, you are spinning so rapidly.  How fast do you spin, exactly?

\par \Merope Well, right now I am spinning roughly 10 times each and every millisecond.  I have become a millisecond pulsar!  

\par \Celaeno Wow!  Such rotational kinetic energy!\footnote{Recall that the more rapid a star is spinning, the larger its total rotational kinetic energy will be, for a given star mass and size.}  Incredible!  And... $\Maia$?  Is that \textit{you}!?

\par \Maia Yes, sister, it is I!  I too have trimmed down quite bit, have I not?

\par \Celaeno You certainly have!  Well, not as much as $\Merope$, but still very impressive.  How fast do you rotate?

\par \Maia Uh... Wait, I calculated this once... I rotate once every... few seconds, I think.  ...Yes, that's right!

\par \Celaeno Oh... I see.  That is quite fast too.

\par \nar $\Maia$'s shoulders slumped, clearly disappointed she had been over-shadowed by $\Merope$ in terms of both size and rotation.\footnote{DISCUSS THE DIFFERENCES BETWEEN NEUTRON STARS OR PULSARS AND WHITE DWARFS, IN TERMS OF THESE TWO PARAMETERS.  ONE IS THE SIZE OF MANHATTAN, THE OTHER IS THE SIZE OF THE EARTH, BLAH BLAH BLAH.}  $\Celaeno$ recognized the look.

\par \Celaeno Your color is truly breath-taking, $\Maia$.  As is your glow.  I am in awe.

\par \Maia Thank you, sister!  I suppose I do glow with a certain... je ne sais quoi.
%NL: Not sure if this part is too sappy, but I have always liked the je ne sais quoid expression.  But, if you didn't, it's likely just me and should be removed.

\par \Celaeno That you do, sister.  That you more than definitely do!

\par \nar $\Celaeno$ smiled warmly at her newly returned sisters; she was beyond grateful for this reunion.  $\Celaeno$ turned toward $\Lacedaemon$, who wore a somewhat timid expression.  Suddenly, $\Celaeno$ realized $\Lacedaemon$ had never met or been introduced to his other sisters.  Introductions were in order!

\par \Celaeno  Where are my manners?  $\Maia$, $\Merope$, it is with great pleasure to introduce to you our new brother, $\Lacedaemon$.  Unfortunately, $\Alcyone$ and $\Taygete$ are no longer with us.  Formed from their merger, our new brother is truly a blessed silver lining to their terrible demise.

\par \Lacedaemon Uh... Hello.  It is lovely to meet you!

\par \Merope And you, brother.

\par \Maia Hello!  I would love to hug my newfound brother but, based on a rough estimate of our current relative velocity, we will have to wait about another billion years or so.

\par \nar $\Lacedaemon$ laughed, delighted by her sister's honesty and sense of humor.

\par \Lacedaemon A very welcome anticipation, my new sister.  Hello to you both.  Wonderful to finally meet you.  I have heard a great deal about you.

\par \Merope Only good things I hope?

\par \Celaeno Absolutely not.  I stuck exclusively to the most terrible things!

\par \nar Joking, $\Celaeno$ smiled and winked at her sisters.

\par \nar Rather suddenly, $\Sterope$ came flying in at several hundreds of kilometers per second, heading from the Galactic Halo.  As you will recall, she was previously ejected by two black holes residing in a dense, massive outer halo globular cluster.  The black holes who assisted her were brothers, named $\Ares$ and $\Zeus$, and had fated themselves to an inevitable merger in order to rescue $\Sterope$ from the gravitational bounds of their host star cluster. \footnote{WAS THIS MADE CLEAR IN THE RELEVANT CHAPTER?!} Drawing closer, $\Sterope$ sees and recognizes her long-lost siblings, and does her best to veer toward them.  She manages to impart a slight gravitational deflection using a nearby giant molecular cloud,\footnote{By passing sufficiently close to a Giant Molecular Cloud (GMC), with a total mass several orders of magnitude larger than that of $\Sterope$, with low relative velocity at closest approach, the impulse imparted to $\Sterope$ by the GMC is maximized.  This means that a maximum deflection is imparted, and $\Sterope$'s trajectory is altered non-negligibly.  Diagram?} and finally passes within earshot of her family.  

\par \Sterope Hello sisters!

\par \nar $\Sterope$ remains within earshot for only a few thousand years, whipping by at high velocity.

\par \Sterope Good bye sisters!

\par \Maia $\Sterope$!  We are all very happy to see you.

\par \Celaeno How we have missed you!  Are you well?

\par \Sterope I do apologize for my high relative velocity.  It is the result of not one by \textit{two} close encounters, each with a particularly kind black hole.  They were brothers, in fact, and their names were $\Ares$ and $\Zeus$.  It seems I am quickly receding from you.  But, yes, I am doing quite well, and have missed you all terribly.  How are you?

\par \nar In unison, the five sisters begin to speak at once, each telling the tale of their own tremendous journey through the Galaxy.  $\Sterope$ furled her brow, trying frantically to decipher all of the updates simultaneously bombarding her.  When they had all finished, miraculously, she was satisfied she had adequately absorbed each and every one of their individual stories.

\par \Sterope You have all experienced so much.  And how you have matured!  

\par \nar $\Sterope$ was now fading in to the distance, forced to say goodbye to her siblings one final time.

\par \Sterope Well, I am off to continue my journey!  Wait!  Have you seen $\Electra$?  Is she well?

\par \Celaeno Yes!  She is over there in the distance, apparently heading toward us from the Galactic Center.  Oh, the stories she will surely tell!  When she arrives, we will extend your most sincere and best wishes.

\par \Sterope Please do!  I love you all!  Take care of yourselves, and each other!

\par \nar With a final twinkle, $\Sterope$ was gone.

\par \Maia How wonderful to see our sister.  She has grown so beautiful over the years.  I am glad she is well.

\par \Merope As am I, sister.  As am I.

\par \Celaeno And I!

\par \Lacedaemon And... I.  I mean, it was wonderful to meet my other sister.  She seemed wonderful.

\par \nar Only a few hundred million years later $\Electra$ finally appears, charging toward her four long-lost siblings at high velocity.  She had been ejected from the Galactic Center by $\Chiron$, a rather helpful super-massive black hole.  

\par \Electra Sisters!  I have found you, at long last.  You look tremendous, all of you!  Oh my, how $\Maia$ and $\Merope$ have changed.  You both look so svelte!

\par \Maia Thank you, thank you.  I feel good.

\par \Merope As do I.  And you look as though you have hardly aged, sister.  You look as young as the day we left you.

\par \Electra Well, I have been exercising.  I've basically been sprinting, apparently toward the four of you, for the last billion or so years.  This is thanks to the help of a super-massive black hole lurking in the heart of our Galaxy!\footnote{IS THIS TIMESCALE RIGHT?  AT SOME POINT, WE NEED TO REVISE THE WHOLE BOOK, WITH AN EYE TOWARD ALL TIMESCALES, DISTANCES AND VELOCITIES PRESENTED, TO MAKE SURE THEY ARE CORRECT/REASONABLE.  THE MAP OF THE SIBLINGS' TRAJECTORIES AARON IS GOING TO MAKE WILL HELP A LOT...}  

\par \Celaeno I have missed you, sister.  Wonderful to see you again.  But, yes, there are only four of us.  You have just missed $\Sterope$.  But do not fear, she is doing well and says hello.

\par \Maia Wait, did you say that you met a \textit{massive} black hole?

\par \Electra I encountered a \textit{super}-massive black hole, in fact!  He was very kind.  His name was $\Chiron$ and he imparted upon me a colossal acceleration.  A gravitational slingshot, they call it.\footnote{Ordinarily, the term "gravitational slingshot" refers to the ejection of a single very high velocity star, during an interaction between an SMBH and an incoming binary star system. In this scenario, to conserve energy, the single escaping star is ejected at high-velocity, so to compensate, its initial binary companion ends up compactly bound to the central SMBH.   More generally, the "gravitational slingshot mechanism" involves three objects, and results in the production of one very high-velocity escaper (typically the least-massive interacting object).  The other two objects could both be SMBHs or IMBHs, or they could be stars.  In general, the more massive is the left-over binary, the higher the final escape velocities of the ejected star.}  That is the origin of my ludicrously high velocity.  Speaking of which, I am now nearly out of earshot from you once again.  How fantastic it was to see you again.  Take care, all of you!  Until we meet again! 

%Suddenly and out of nowhere, $\Merope$ and $\Maia$ appear as well.  They are no longer recognizable, but they of course recognize their siblings.  So they must re-introduce themselves, but their siblings eventually recognize them, and are happy to see the,

\par \nar $\Electra$ now gone, $\Lacedaemon$ belched without warning.
%NL: I don't know if the above works.  Remove?  Probably too many fart/burp jokes...  Sigh, I'm such a child.

\par \Lacedaemon Sorry about that.  Residual nervousness...

\par \Celaeno Uh... Well, I was going to say that it was wonderful to see $\Electra$.  She seems happy, and excited for her next adventure.

\par \Maia She did seem happy, didn't she?  

\par \Merope A true family reunion!  I could not be happier.

\par \nar But before $\Merope$ could finish, $\Celaeno$ and $\Lacedaemon$ had drifted beyond earshot of $\Maia$ and $\Merope$.  Their relative kinetic energy with respect to each other was sufficiently high, after all, that they were not gravitationally bound.%\footnote{I BELIEVE WE DISCUSSED BINARY BINDING ENERGY IN A PREVIOUS CHAPTER, BUT THIS CONCEPT SHOULD BE EXPLAINED USING THE PRIOR DEFINITION AND REFERRING TO IT.}  
Their inevitable fate, complete dissociation, was upon them.

\par \Lacedaemon What was that, sisters?  We can hardly hear you anymore!

\par \Celaeno It seems our time together has once again come to an end.  At least for now.  Fare thee well, sisters!  We shall carry you with us in our hearts always.  Until we meet again!

\par \Maia And us, you, my sister!

\par \Merope We shall remember you with fondness each and every day.  Until the next time!  

\par \nar The entire family disperses yet again.  This time, they are reassured by the fact that, given enough time (which is plentiful in the life of a star) they are bound to meet again in the distant future.\footnote{This isn't even remotely true.  While the exact number is tricky to calculate, the probability that six or seven stars dispersed randomly across the Galaxy will find themselves all within a small volume at the same time is effectively zero over the lifetime of the Universe.}

%NL: NEEDS A BETTER ENDING!!!

\end{document}
