\documentclass[main.tex]{subfiles}
\begin{document}

\chapter{The Making of a Millisecond Pulsar...}

\par \nar \rmmerope~ was now a proud neutron star.\footnote{Neutron stars (NSs) are stellar remnants; the evolutionary descendants of their once proud progenitor stars, post-supernova explosion.  These are massive stars with total masses above $\gtrsim$ 8 M$_{\odot}$.  There is a poorly known upper mass limit for NS formation, indicating the cross-over between collapse to a neutron star or to a black hole.  This critical progenitor mass is thought to lie around 40 M$_{\odot}$, but the exact number is highly uncertain both theoretically and empirically.}   But she was a little too close to \rmmaia~ for comfort; plasma was being pulled from \rmmaia's equator, flowing toward \rmmerope~ and gathering around her equator as an accretion disk.\footnote{Accretion disks are, even in their simplest forms, remarkably complicated structures.  Many disk parameters (e.g., density, scale height, opacity to radiation, etc.) are important in deciding their subsequent time evolution.  Due to conservation of angular momentum, they tend to begin as thin disks of gas orbiting the accretor.  Depending on how the disk responds to radiation from the host object, it can become puffed up at various distances from the central orbiter. Ultimately, gas particles in the disk collide into each other, viscously redistributing angular momentum. This causes some particles to end up with larger angular momenta and diffuse outward to larger disk radii, and some particles to end up with lower angular momenta and diffuse inward.  Inward diffusion is the mechanism responsible for direct accretion of gas particles onto the central object, and their removal from the accretion disk. CAN WE INCLUDE AN ILLUSTRATION OR SCHEMATIC DIAGRAM?}  Viscosity caused the disk to diffuse inward until, from the inner boundary of the accretion disk, material began to flow and accrete on to \rmmerope's surface.\footnote{Viscosity is a parameter used in fluid dynamics that characterizes the degree of internal friction of a given medium.  Viscous substances tend to diffuse very slowly.  For example, water and honey have low and high viscosities, respectively.  The underlying mechanism responsible for viscosity can vary, but ultimately stems from inter-particle forces.  CHECK!  The viscosity of the gas or accreted material ultimately decides the rate at which low angular momentum gas particles are produced and accreted onto the central object.  More viscous substances have longer diffusion times.}  The material brought in angular momentum, causing \rmmerope's spin rate to slowly increase over time.\footnote{Although it tends to be low angular momentum material (relative to the integrated distribution of, or total, angular momentum in the disk) that is accreted from the disk onto the central object, the particles nevertheless carry non-negligible angular momentum.  Hence, to conserve angular momentum during the accretion process, orbital angular momentum in the disk tends to be converted into spin angular momentum in the central accretor.  in other words, the accretor is "spun up", rotating at ever higher rates.  In the case of millisecond pulsars, the neutron stars rotate once every millisecond!}

\section{Spun.  Fully spun.}

\par \Merope Whoa, I am so sorry about this.  I did not mean to accrete from you.

\par \Maia It is not your fault, sister.  The sudden change in our orbital parameters due to your supernova explosion caused it.\footnote{REFER TO WHERE ORBITAL PARAMETERS ARE DEFINED IN AN EARLIER CHAPTER.}  You, of course, had no say in the matter.

\par \Merope I did not.  Nevertheless, I cannot help but wish that I were smarter, and could have thought of some way to deliver us to a less dramatic outcome.

\par \Maia Oh, my sister.  I do love you for your tenacity.  But you must not be so hard on yourself.  It is not your fault, and you must let it go.  After all, I feel fine.  I am slowly losing mass to you, but the consequences seem quite benign so far.  How are you doing?

\par \Merope Oh, alright, I suppose.  I am rotating ever faster.  On the one hand, it is making me \textit{very} dizzy.  \textit{But}, on the other hand, I can feel the excess energy stored within me.  It feels like...power.  I sort of like it.  ...Should I be ashamed?\footnote{Significant rotation rates tend to come along with significant rotational kinetic energy (and, apparently shame), but it ultimately depends not just on the spin rate but also the properties of the rotating object.  EXPLAIN THE CONCEPT OF ROTATIONAL KINETIC ENERGY, AND MOMENT OF INERTIA.}

\par \Maia No, I do not think so.  Power can be spent for good, and the prospect of doing good should indeed make you feel warm inside.  So, in the end, it is up to you how you decide to spend your new-found power.

\par \Merope  Well, to be honest, it seems at the moment like storing it away is the best thing for us all.  But I do hope that the day will come that my new-found rotational energy will prove critical to improving the state of the Universe.

\par \Maia And that is one of the many reasons why I love you, sister.  

\par \Merope Uh... I think I am rotating...carry the 1...about once every...uh...millisecond... Can that even be right?
%NL: Does the above ``joke'' work?

\par \Maia Let me see.

\par \nar \rmmaia~ watched intently as \rmmerope~ rotated.  A keen eye, she was able to count the number of rotations over a given minute, and from there calculate the rotation rate per second.\footnote{To calculate the rotation rate in this way, \rmmaia~  simply counted the number of complete revolutions performed by \rmmerope~ over some fixed interval of time.  The rotation period is then the number of revolutions divided by the total time interval (in this case one minute).}

\par \Maia Holy cow!  You are right! You rotate roughly once every 10 milliseconds!  That is \textit{very} fast.  

\par \Merope Wow.  It feels like I am rotating quickly, but hearing the exact number really drives it home.  Okay, well, I'll be over here waiting for something to impart all this rotational kinetic energy to, hopefully for the greater good. 

\par \nar \rmmerope continued to accrete from \rmmaia, although most of the infalling matter was now being expelled via the ``propeller effect''; \rmmerope now possessed strong magnetic fields, that repelled incoming charged particles.  As \rmmaia lost mass, the distance separating her from \rmmerope grew; their orbital separation was increasing.\footnote{DO A SIMPLE CALCULATION USING CONSERVATION OF ENERGY AND ANGULAR MOMENTUM TO ILLUSTRATE THIS.}  Regardless, \rmmaia had a large enough radius that she continued to fill her Roche lobe\footnote{The Roche lobe defines the critical distance from the center of a star's mass beyond which gravity will pull matter away from the star's surface and on to its binary companion. DIAGRAM?}, steadily losing mass to \rmmerope.  Eventually, \rmmaia had lost so much mass she hardly felt like herself anymore.

\section{Life After Death...}

\par \Maia Wow, I have been widdled away, sister!  I am but a shadow of my former self.  I must have lost most of my original mass by now.  

\par \Merope I am sorry, sister.  How do you feel?

\par \Maia Hmmm... Fine considering, I suppose.  I am starting to sweat a little bit though.  Do I look hot at the surface?

\par \Merope Now that you mention it, you are glowing somewhat.  And your color has changed. You're much whiter than before.

\par \Maia Hmmm...  Well, this is embarrassing.  I feel as though my core has been exposed.\footnote{In the cores of evolved stars, and inert ash of heavy nuclei sit at the bottom of the star's inner potential well.  For high-mass stars, these nuclei can include helium, carbon, oxygen, silicon, magnesium, etc.  Once the outer envelope of hydrogen-rich plasma is accreted from such an evolved star, the inner core is left exposed.  It is much hotter than the surface layers of its progenitor star, and so shines a whiteish blue and emits significant ultraviolet radiation.}    

\par \nar \rmmerope~ closed her eyes, turning away from \rmmaia.

\par \Merope Don't worry, sister.  I am not looking.  Your privacy is safe with me.

\par \Maia Well, that is very considerate.  But... I do not see how to cover up, now that my envelope is gone. I fear you will have to accept me as I now am.

\par \nar \rmmerope~ turned back toward her sister, a smile spread across her face.

\par \Merope I will take you in any form, sister.  I am just glad that you are okay.  How do you feel?
%NL Is this chapter perhaps a little cheesy?  If so, someone want to take a crack at editing accordingly?

\par \Maia Well, all things considered, I feel okay.  Hotter at the urface, a lot more compact, certainly less luminous, but otherwise fine.\footnote{REVIEW HOW ALL THIS TIES IN TO THE PROPERTIES OF WHITE DWARFS.}  How do I look?

\par \nar \rmmaia~ put her hands on her hips, and swiveled back and forth.  Glamorous did not do her justice.

\par \Merope Wonderful!  You have never been more symmetric!  I am truly jealous!

\par \Maia Oh, please, your symmetry is perfect.  

\par \Merope Not with all this rotation.  Look at \textit{this}.

\par \nar \rmmerope~ pointed angrily at her equator. 

\par \Merope I have a bulge!  A bulge!

\par \Maia Sister, you are still the most beautiful of all the Universe's creatures.
%NL: Yep, definitely got cheesy...  Dammit.

\par \Merope There is no doubt that you are my sister.  And even less doubt that I love you tremendously.

%EDIT THE BELOW!!!
\par \nar \rmmaia~ and \rmmerope, out of reach, extended open arms toward each other.  Clawing futilely at vacuum in each other's general direction for but a few seconds, they quickly exchanged a mutual glance of loving understanding, and then ceased their frantic flailing.

\end{document}
