\documentclass[main.tex]{subfiles}
\begin{document}

\chapter{Taming the Beast}

%\MAIA AND \MEROPE ARE IN A BINARY!!!  NEED TO INCLUDE \MAIA IN THE NEXT CHAPTER!!!

\par \nar \rmelectra~'s journey began when she left her birth cluster, heading out of the Galactic Disk.  But it wasn't long before a close encounter with a giant molecular cloud, named \rmaethra~, diverted her course.  The gravitational pull exerted by the cloud caused a nearly 90$^{\circ}$ deflection.\footnote{INCLUDE DIAGRAM SHOWING THE EVOLUTION OF THE INTERACTION.}  

\section{A Change of Course...}

\par \Aethra I am so sorry, child.  It is often that I under-estimate the strength of my own gravity.  It would seem I have diverted your course through the Galaxy.

\par \Electra Oh, that's quite alright.  Honestly, it could have happened to anyone.

\par \Aethra It is true, but in this particular case, I have inadvertently diverted your trajectory directly toward the Galactic Center.

\par \Electra Well, that sounds mildly terrifying.  I \textit{was} feeling a little anxious about the uncertainty in my new trajectory, and my new ignorance regarding where exactly in this wilderness I will eventually find myself.  I'm afraid to even look...

\par \nar Turning in the direction of her bulk motion, \rmelectra~ suddenly realized that, as \rmaethra~ had pointed out, her new course through the Galaxy pointed right toward the heart of the Galaxy (i.e., the Galactic Center).\footnote{DESCRIBE THE CENTRAL NUCLEAR STAR CLUSTER IN THE MW.}

\par \Electra Uh oh...  That thing off in the distance looks to be a very dense cluster of stars and gas.  It looks to be a messy environment too.  Well, much messier than around these parts, at least, since mostly it is empty space punctuated by the odd star.

\par \Aethra That's correct, child.  In the Galactic Center, there lives a dense nuclear star cluster, jam-packed with stars spanning all sorts of masses and ages.  We are in the field of our Galaxy currently, where the mean stellar density is about 1 star per cubic parsec.\footnote{Said another way, the distance separating the infamous Sun and her next nearest neighbor, Proxima Centauri, is about a parsec.  But if you take the Sun and Proxima Centauri as they are now, and drop them in to the very central regions of the nuclear cluster at the heart of our Galaxy, then more than 100 stars will fall between the pair.  The stellar densities are that much higher!}

\par \Electra Wow!  That is dense!  I guess I will find out for myself when I get there.

\par \Aethra Rest assured, young one, anything could still happen.  But, if you aren't careful, you stand to meet the behemoth that lurks at the Galactic Center.

\par \Electra Wait.  What's a behemoth?

\par \Aethra In this case, it is a super-massive black hole, weighing in at well over a million solar masses, that lives in the central nuclear star cluster surrounding the very heart of our Galaxy.\footnote{GIVE SOME BACKGROUND ABOUT WHAT WE KNOW ABOUT SMBHS - THEIR TYPICAL MASSES, SIZES, HOW THEY FORM, OBSERVATIONAL PROOF OF THEIR EXISTENCE, ETC.}  

\par \Electra A super-massive black hole (SMBH)?!?  What's their story?

\par \Aethra They are dark, and do not emit light.  And they are very dense; typically smaller than a planet in diameter.  You will only ever be aware of them from their gravitational influence.  For example, if you drift within the central parsec\footnote{MAKE SURE THE TERM PARSEC IS DEFINED UPON ITS EARLIEST APPEARANCE.} or so of the central nuclear star cluster, you will start to feel the gravitational influence of the central SMBH non-negligibly.  That is, your trajectory will start to be dictated almost entirely by the gravity of the SMBH.  The total masses of SMBHs in the Universe can range anywhere from 10$^6$ to 10$^{10}$ times the mass of a typical star.\footnote{Typically, we define astronomical masses this large to be in units of the mass of the Sun.  However, since there is no a priori reason for \rmelectra~ or \rmaethra~ to be aware of our Sun, and the Sun is remarkably representative of the average or typical star in our Galaxy, we will define units of mass with respect to a "typical" star or, equivalently, the Sun.}  In our own Milky Way, the super-massive black hole is only about 4 $\times$ 10$^6$ times the mass of a typical star, so it's rather puny compared to some of the other SMBHs in the Universe!

\par \Electra Okay, that sounds like a lot of power.  A Destroyer of Worlds.  A Deliverer of Death.  

\par \Aethra It is.  If stars drift too close to the central behemoth, they can be torn apart.  Their corpses are briefly crushed upon approaching very close to the SMBH, and they become extremely luminous for a brief time.  Some of their corpse is eaten by the behemoth, and some escapes as high-velocity ejecta; the tenuous final remains of a once proud star.  And, once inside an SMBH's belly, it is impossible to escape!  The very laws of the Universe forbid it! \footnote{DEFINE A TIDAL DISRUPTION EVENT AND PROVIDE SOME BACKGROUND.}

\par \Electra Oooookay.  Gotcha.  Stay the heck away from the SMBH lurking at the heart of the Galactic Center.  I will keep my eyes peeled for it.  Oh wait....it's invisible.  Well, I guess I will just have to keep my wits about me.

\par \Aethra I am glad, child, and wish you the best of luck.  May the Cosmos bestow great luck upon you.

\par \Electra That sounds pretty great, actually.  I'll take two!
%NL: Does this joke work?

\section{Journey to the Heart of the Galaxy} \label{journey}

\par \nar \rmelectra~ continued her inevitable journey toward the Galactic Center, now too far from \rmaethra~ to see or hear her.  Along the way, \rmelectra~ met a number of colorful characters.  The first of these, \rmdardanus, took a particular interest in understanding her trajectory through the Galaxy.  She came upon him while traveling through the low-density Galactic field, No Man's Land, slowly drifting by at a distance, but close enough that they could hear each other bellow from a distance.

\par \Dardanus Hello there!  My name is \rmdardanus.  Where are you off to in such a hurry?!

\par \Electra  Hello!  I am \rmelectra.  Well, if you take my current trajectory through the Galaxy and propagate it forward through time, it looks to me like I am ultimately heading for the Galactic Center.

\par \Dardanus  Great!  Wait... What is a ``Galactic Center''?

\par \Electra  As far as I know, it is the very Heart of our Galaxy.  There, a dense nuclear star cluster lives, where the stellar densities are extremely high.  If I have my math right, it is about a million times denser in nuclear star clusters than it is here!
%Here, a dense nuclear star cluster lives, jam-packed with stars spanning all sorts of masses and ages.  We are in the field of our Galaxy currently, where the mean stellar density is about 1 star per cubic parsec.  Said another way, the distance separating the infamous Sun and her next nearest neighbor, Proxima Centauri, is about a parsec.  But if you take the Sun and Proxima Centauri as they are now, and drop them in to the very central regions of the nuclear cluster at the heart of our Galaxy, then more than 100 stars will fall between the pair.  The stellar densities are that much higher!

\par \Dardanus  Wow!  That sounds crowded.  I bet you can hear \textit{everything} \textit{everyone} is doing or saying, at any given moment.  

\par \nar \rmdardanus~ shudders at the thought of living in an environment devoid of privacy, having grown used to the relative isolation characteristic of No Man's Land; it has been home to him for many billions of years now.

\par \Electra But the really terrifying thing, if you ask me, is the super-massive black hole lurking in the central nuclear star cluster.

\par \Dardanus A super-massive black hole, you say?  What is that?

\par \Electra It is a dense dark object that does not emit light.  If I remember right, they have total masses that can range anywhere from 10$^6$ to 10$^{10}$ times the mass of a star such as yourself.  In our own Milky Way, the super-massive black hole is something like 10$^6$ times the mass of a typical star.  So it's massive, but not compared to the rest of its siblings.

\par \Dardanus Holy cow!  That's a lot of mass.  In fact, couldn't things potentially get violent if you were to wander too close to an SMBH?

\par \Electra Uh... Yeah, it's very possible.  Technically, if you wander too close to a super-massive black hole, it can eat you whole.  And, once inside its belly, you can never escape.

\par \Dardanus  Oh, wonderful!  That doesn't sound even remotely terrifying...

\par \Electra That's what I've been saying!

\par \nar \rmdardanus~ was no stranger to sarcasm.

\par \Electra Um... And even if you don't get eaten, if you \textit{do} wander close to a super-massive black hole, you are almost certainly going to get accelerated to extremely high velocities.  Like, we are talking thousands of kilometers per second.

\par \Dardanus  WHAT?!  Is that even possible?

\par \Electra  Yep, such fast stars have definitely been seen whipping through our very own Galaxy, presumably on their way out.  I know, I've seen a couple since leaving my birth cluster.

\par \Dardanus Well, I said it once and I'll say it again:  this all sounds very terrifying. You're definitely going to need to keep your wits about you in there.

\par \Electra Exactly!  I mean, given all these hurdles, how can I possibly expect to adequately prepare myself for a journey through the central nuclear star cluster of the Milky Way?  What I'm trying to say is:  I'm definitely going to die in there!

\par \nar \rmdardanus~ yawns, tired from a rare conversation.  

\par \Electra  Do you even hear me?  \textit{I am going to die for sure!}  And there's nothing I can even do about it.

\par \Dardanus Well, if I were you, the first thing I would do is take another nap, frankly.  I figure you'll need to be fresh and alert for the end of the long journey to get yourself to the Heart of the Galaxy.

\par \Electra SLEEP?!?  Who has time for sleep?  I need to craft an ingenious survival plan, gather supplies, panic a few more times, etc.  You get the idea.  
%Oh my gawd, you don't have a plan!  You are so going to die!  And why am I the only one freaking out about this?

\par \rmdardanus~ had already fallen asleep, now snoring loudly.

\par \nar Long after parting ways with \rmdardanus, \rmelectra~ too falls asleep...although the frequent and violent nervous twitching appears anything but peaceful.

\section{Befriending the Beast} \label{beast}

\par \nar \rmelectra~ awoke surrounded by stars.  This definitely was not the Galactic field anymore.  Recalling her trajectory through No Man's Land before falling asleep, \rmelectra~ suddenly realized where she was.

\par \Electra The Galactic Center!  I made it!  Hello there!  Can you tell me:  Is this the nuclear star cluster at the heart of the Milky Way?

\par \nar A nearby star named \rmcarystus~ turned to face \rmelectra, the confusion evident on his old face.  

\par \Carystus Well, to be honest, I've lived here my entire life, and I really have no idea what stars in the rest of the Galaxy call this place.  But, what I can tell you about this unique place in the Galaxy, is that you will not find a more extreme population density anywhere else in the entire Milky Way.  We pride ourselves on our crazy stellar densities, reaching as high as $\sim$ 10$^7$ stars per cubic parsec!

\par \Electra, Yep, this \textit{must} be the Galactic Center I have been looking for.  The behemoth must be close!

\par \Carystus Behemoth?

\par \Electra Allegedly, a super-massive black hole resides here somewhere, hiding among its many luminous denizens.

\par \Carystus Oh!  You must mean \rmchiron.  He's here alright.  Knowing when you are close is one thing, but finding him without being eaten is altogether another. 

\par \Electra Oh wow!  He sounds dangerous but illusive.  Just like he was described to me!  Any tips for how best to approach him?  I'd very much like to meet and interact with him... I mean, from a safe distance.

\par \Carystus  You are a brave one!  That is a good question.  Hmmmmm...   Well, \rmchiron~ is surrounded by a dense population of massive stars, directly orbiting his significant center of mass.\footnote{DESCRIBE WHAT IS KNOWN FROM OBSERVATIONS ABOUT THE INNER ~0.1 PC OF THE MW.}  These stars all emit winds from their surfaces, at a rate of many thousands of times the mass of a typical star per year.\footnote{Describe mass loss in massive stars and describe their basic evolution compared to low-mass stars.}  This is fast enough that, within almost no time, \rmchiron~ will collect their expelled gas in to a compact disk, which he accretes from and is thus his main source of fuel.  This disk, once formed, should emit light and be luminous; it is directly visible to us all on its own, even if \rmchiron~ is not.  But, without a doubt, \rmchiron~ will reside at the center of any such disk, called an ``accretion disk''.  This is where you can be sure to find him.

\par \Electra  Look for a small but bright accretion disk at the very center of this nuclear cluster.  This will clearly and unambiguously identify \rmchiron~.  Gotcha!  

\par \Carystus  Yep, that's about it.  If you can do that, you'll behold the beast.  

\par \Electra That's the goal!

\par \Carystus Just be careful.  Super-massive black holes have a mind of their own, and can be unpredictable.  Strange physics you might not be used to can emerge and even dominate the picture in this new high-gravity regime.\footnote{BRIEF INTRO TO GENERAL RELATIVITY!}  When these effects become important can be hard to identify before they are right on top of you.  So one must take great care to safely navigate such an extreme and dangerous environment.

\par \Electra I read you, loud and clear.  I definitely intend to keep my wits about me.  Well, I'm off to meet \rmchiron.  Wish me luck!

\par \Carystus Good luck, my child.  May the brightest of futures befall you.

\end{document}
