\documentclass[main.tex]{subfiles}
\begin{document}

\chapter{A Tale of Two Black Holes...}

\par \nar \rmsterope~ was trapped in an ancient globular cluster and in search of a pair of black holes.  The idea was to first encounter a binary star system, undergo a chaotic three-body interaction, and ultimately become exchanged into it.  From there, she need to find a black hole, undergo another a chaotic three-body interaction, and eject her former binary companion.  Finally, she needed to encounter a second black hole and undergo yet another chaotic three-body interaction, this time with two black holes.  In the end, she hopes to be the one ejected, leaving the two black holes behind in a binary and escaping from the cluster altogether.\footnote{DESCRIBE THIS PROCESS IN DETAIL, AND DICUSS ITS RELEVANCE FOR CHAOS.  PERHAPS SHOW THE HUT \& BAHCALL FIGURE OF THIS?}  

%\section{A Tale of Two Black Holes}

\par \nar She wandered through her new and hopefully temporary home for many millions of years.  During this time, she endured many encounters with the other denizens of the cluster.  But it took several such interactions before she eventually met even one black hole.  Finally, she stumbled upon a suitable target.  \rmsterope~ had managed to exchange herself in to a normal stellar binary by the time she encountered the first black hole.  As the black hole drew close, its gravity pulled \rmsterope~ and her companion toward it.  A chaotic three-body interaction ensued.  A delicate dance, the three bodies bobbed and weaved by each other, continually accelerated by each other's mutual gravitational attraction.  They took turns being flung out far from the system center of mass, on loosely bound excursions.  All this, but without imparting enough kinetic energy for any one of them to escape, each excursion carefully choreographed such that they remained bound, destined to return to rejoin the melee.  The dance continued until, eventually, \rmsterope's original companion was ejected, flung off at sufficiently high velocity to escape to what amounts to spatial infinity in \rmsterope's reference frame.\footnote{DISCUSS THE CONCEPT OF SPATIAL INFINITY, AND REFERENCE FRAMES.}

\section{A New Companion...}

\par \Sterope Uh... Hello there... Wherever you are.  I can't actually see you, you being completely black and all.  But I know you're there!  I must be orbiting \textit{something}, after all.

\par \Ares You are not alone.  It is I, \rmares.  Star destroyer.  Devourer of worlds.

\par \Sterope That sounds scary.  Regardless, I'm pleased to meet you.  I've been searching for a black hole such as yourself for quite some time.

\par \nar \rmsterope~ extends an eager hand, hoping for a handshake but suddenly realizing \rmares~ could easily be out of reach.\footnote{AGAIN, DO WE WANT THE STARS TO HAVE HANDS?}

\par \Ares You aren't scared?

\par \Sterope Scared of what?

\par \Ares Me!  I just said I destroy stars!

\par \Sterope Oh.  Right.  Well I wish you the best of luck with that, but so far you seem nice enough.

\par \nar \rmares~ lets out a long sigh.

\par \Ares All this mass, and nobody around here seems to care.  What gives?

\par \Sterope I don't know, but \textit{I} sure care.  That's why I came looking for you.  I need an object of significant mass to help me escape from this cluster.  Well, two massive objects, to be precise.  

\par \Ares Ah, you are in need of a gravitational assist!  A slingshot from a massive body.  Finally, a use for me!\footnote{DEFINE THE TERM GRAVITATIONAL ASSIST.}

\par \Sterope Yes!  You could be very useful to me.  

\par \Ares Wonderful.  I feel useful already.

\par \Sterope Okay, so, how do we do this?

\par \Ares Well, I guess I could take a run toward you, and maybe that would work?

\par \Sterope I don't think so.  We are gravitationally bound.  I do believe we need at least some form of outside interference to unbind us.  After all, the excess energy must come from somewhere.\footnote{REMIND THE READER ABOUT ENERGY CONSERVATION AND ORBITAL BINDING ENERGY.}

\par \Ares Fair enough.  We need another black hole!

\par \Sterope Exactly!  I am confident that, between the two of us, we can find one.

\par \Ares I might know where.  My brother is lurking around here somewhere. 

\par \Sterope You're brother?!

\par \Ares \rmzeus~! \rmzeus~!  Where are you, brother?

\par \nar A long silence passed.  \rmsterope~, excited at the prospect of a second black hole, listened closely for a response.  Finally, a response came.

\par \Zeus I am here.   

\par \Ares  Wonderful.  Good to hear your voice again, brother.

\par \Zeus And you, brother.

\par \Ares I have met a friend, \rmsterope~, who requires our assistance.

\par \Zeus  Hello, \rmsterope~.  Nice to meet you.

\par \Sterope And you, \rmzeus~.  Truly, this is intimidating.  You are both \textit{so} massive.

\par \Zeus Do not be intimidated, child.  Our mass can at best accelerate you, nothing more.

\par \Sterope You are both so kind.  I don't understand why everyone I met before you was so afraid of black holes.

\par \Zeus A common misconception. 

\par \Ares \rmsterope~ is in need of a gravity assist, to escape from this cluster.

\par \Zeus I see.  Well, you seem kind, noble and out of place here amongst these old stars.  I am happy to offer my mass in the name of this cause.  But how do you propose we get the job done?

\par \Ares As god of war, I have facilitated many a squirmish in this cluster environment.\footnote{USE N$\sigma$V ARGUMENTS TO EXPLAIN THE RATE OF SUCH INTERACTIONS.}  As I see it, there is no better way to impart a velocity kick of sufficient magnitude to eject \rmsterope~ from the cluster.  Us three must engage in a chaotic dance, choreographed by none other than gravity.

\par \Zeus  Very well.  I am coming toward you.  Prepare yourselves for chaos!

\par \nar \rmzeus~ drifted ever closer to \rmsterope~ and \rmares~, slowly at first but rapidly gaining speed.  Before they knew it, \rmzeus~ was upon them, tugging them in every which direction, and vice versa.  Chaos ensued.  The two black holes and \rmmerope~ completed a series of wild loops and orbits.\footnote{A SKETCH OR DIAGRAM DEPICTING THIS WOULD BE COOL AND USEFUL.  AARON?}  More than a few times, \rmmerope~ passed close enough to one of the black holes to really feel their gravitational might.  The differential force of gravity caused matter to be pulled from her surface after making her closest approach.\footnote{Consider a normal stellar binary star system, composed of two stars.  Recall that, due to tidal forces, the side of a finite-sized object closest to its binary companion will be more strongly attracted by gravity than the opposing side, raising a "tidal bulge".  If the radius of one star exceeds the "Roche limit", then mass is transferred from the surface of the star onto the surface of its binary companion.  The physics here is ultimately the same, but the dance is chaotic; this scenario occurs sporadically and pairwise over the course of the chaotic three-body interaction.} Not much matter, but enough to freak out both her and her interloping black hole; each time this occurred, they exchanged a startled glance.  Finally, \rmmerope~ was ejected at high velocity, leaving \rmares~ and \rmzeus~ bound in a new binary.    

\par \Ares I think that should do it.  

\par \Zeus Agreed!  Surely she must be flying away at above the escape velocity.  

\par \Ares Fare thee well, \rmsterope~!  May only glorious adventures befall you!

\par \Sterope Thanks so much, to the both you!  Free at last!

\par \nar Sure enough, \rmsterope~ escaped from the gravitational bounds of her temporary host cluster with kinetic energy to spare.  Her trajectory now pointed inward, back toward the Galactic Disk.  Optimistic she would find her lost siblings, \rmsterope~ smiled at the thought of seeing them again..

\end{document}






